---
title: "R Assignment 2"
author: "Ashtami Bhuleskar 18201912"
date: "November 14, 2018"
output:
  pdf_document: default
  html_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> **Data Programming with R**

#Task 1: Analysis#

##1.1: Load the lawyers' data into R. What proportion of the lawyers practices litigation law?##

```{r}
#Loading file
LawyerData <- read.csv("C:/Users/Ashtami/Documents/R/Assignment 2/Lawyers.csv", stringsAsFactors = F)
LawyerData <- data.frame(LawyerData)

#Task 1 : 1
#Using prop function to find proportion
Litigation <- round(prop.table(table(LawyerData$Practice)),2)[2]
print(Litigation)
```

##1.2: Is the proportion of lawyers in the Boston office that practice corporate law higher than the proportion of lawyers in the Providence office that practice corporate law?##

```{r}
#Using if else to check if the value is Higher or not
if(round(prop.table(table(LawyerData$`Practice`, LawyerData$Office)),2)[1][1]>round(prop.table(table(LawyerData$`Practice`, LawyerData$Office)),2)[3][1]){
  print("Higher")
} else {
  print("Lower")
}
```

##1.3: Use the aggregate function to compute the average age of lawyers who practice corporate law and of lawyers who practice litigation law, across the different levels of seniority. Label the columns of the resulting data frame appropriately.##

```{r}
averageAge <- aggregate(LawyerData$Age,list(LawyerData$Seniority,LawyerData$Practice), mean)

#Labeling the columns of the result obtained using aggregate
names(averageAge)[1] <- c("Seniority")
names(averageAge)[2] <- c("Practice")
names(averageAge)[3] <- c("Average Age")
print(averageAge)                       #Printing result
```


##1.4: Which office has the youngest median age?##
```{r}
#Finding median age in all offices
youngest <- aggregate(LawyerData$Age,list(LawyerData$Office), median)

#Labeling the columns of the resultant dataframe
names(youngest)[1] <- c("Office")
names(youngest)[2] <- c("Median Age")
print(youngest)                         #Printing result

#Finding the youngest median age among the resut obtained using aggregate 
min_office <- youngest$Office[youngest$`Median Age` == min(youngest$`Median Age`)]
print(min_office)

test <- names(which.min(by(LawyerData$Age,LawyerData$Office, median)))

```

#Task 2: Writing your own functions#

##2.1: Write a function which compute the Rosenbrock banana function using a loop. Test the function on the vectors x = (:2; :5) and x = (:2; :5; :1; :6)##

```{r}
#Function Rosenbrock Banana using loops
Rbanana <- function(x) {              #Function definition
  s <- vector(length = length(x)-1)   #Creating empty vector of size length(x)-1
  for(i in 1:length(x)-1){            #For loop runs for length(x)-1 times
    s[i] <- ((100 * (x[i+1] - x[i]^2)^2) + (1 - x[i])^2)  #Every element of vector s stores value computed by Rbanana function in the current loop
  }
  return(sum(s))                      #Return cumulative sum of all elements in vector s
}

vector1 <- c(0.2,0.5)                 #Initialize global variable to store first vector of size 2
vector2 <- c(0.2,0.5,0.1,0.6)         #Initialize global variable to store second vector of size 4

banana1 <- Rbanana(vector1)           #Function call to Rbanana function by passing vector 1 as parameter
print(banana1)

banana2 <- Rbanana(vector2)           #Function call to Rbanana function by passing vector 2 as parameter
print(banana2)
```

##2.2: Propose an alternative function that does not use any loop. Test the function on the same two vectors.##

```{r}

#Function Rosenbrock Banana without using loops
Rbanana2 <- function(x){              #Function definition
  first <- x[1:(length(x)-1)]         #Variable 'first' is a list that stores all values of data except the last (1 to second last)
  second <- x[2:length(x)]            #Variable 'second' is a list that stores all values of data except the first(2 to last)
  equation <- sum((100*(second - first^2)^2) + (1 - first)^2)   #Computation equation, sum function iterates over the entire list
  
  return(equation)                    #Returns cumulative sum obtained
}

banana3 <- Rbanana2(vector1)          #Function call to Rbanana2 function by passing vector 1 as parameter
print(banana3)

banana4 <- Rbanana2(vector2)          #Function call to Rbanana function by passing vector 2 as parameter
print(banana4)
```

##2.3: Compare the timings you obtain by repeating the function calls 100 times using the vector x = (:2; :5; :1; :6) as input.##

```{r}

startpoint <- proc.time()             #proc.time to define start point
for (i in 1:100) Rbanana(vector2)     #for loop to call Rbanana function 100 times
proc.time() - startpoint              #End - start to give intermediate time span

startpoint <- proc.time()             #proc.time to define start point
for (i in 1:100) Rbanana2(vector2)    #for loop to call Rbanana2 function 100 times
proc.time() - startpoint              #End - start to give intermediate time span
```


#Task 3: Writing S3 methods#
##3.1: Load in the data as an object called DublinAirport. Assign to the DublinAiport object the classes WeatherData and data.frame.##

```{r}
#Loading data from CSV file
DublinAirport <- read.csv("C:/Users/Ashtami/Documents/R/Assignment 2/2018_09_Dublin_Airport.csv", stringsAsFactors = F)

#Assigning object of DublinAirport to WeatherData and data.frame classes
class(DublinAirport) <- c('WeatherData', 'data.frame')
```

##3.2: Write an S3 summary method for an object of class WeatherData which produces the following statistical summaries for the rain, maxtp, mintp variables: mean, standard deviation, minimum, maximum.##

```{r}
summary.WeatherData <- function(data){  #Defining new summary function under WeatherData class
  
  #Using lapply function on every columns to compute mean, sd, min and max and storing in a list
  mean <- lapply(data[2:4], mean)         
  SD <- lapply(data[2:4], sd)
  minimum <- lapply(data[2:4], min)
  maximum <- lapply(data[2:4], max)
  
  #Creating list of lists to store above created lists
  Summary <- list(mean, SD,minimum,maximum)
  #Renaming for better identification
  names(Summary) <- c("Mean", "Standard Diviation","Minimum","Maximum")
  
  # cat and print functions to print the summary
  cat("MEAN : ", "\n")
  print(Summary[1])
  
  cat("STANDARD DIVIATION : ", "\n")
  print(Summary[2])
  
  cat("MINIMUM : ", "\n")
  print(Summary[3])
  
  cat("MAXIMUM : ", "\n")
  print(Summary[4])
  
}

summary.WeatherData(DublinAirport)      #Calling summary.WeatherData function
```

##3.3: Download the new data set 2018 09 Cork Airport.csv from Blackboard, assign the classes WeatherData and data.frame to the object containing the Cork data, and test your function on it. Interpret your findings for Dublin and Cork Airports.##

```{r}
#Loading Cork data from CSV
CorkAirport <- read.csv("C:/Users/Ashtami/Documents/R/Assignment 2/2018_09_Cork_Airport.csv", stringsAsFactors = F)

#Assigning object of DublinAirport to WeatherData and data.frame classes
class(CorkAirport) <- c('WeatherData', 'data.frame')

#Calling summary.WeatherData function
summary.WeatherData(CorkAirport)
```
#####Interpreted findings########

Maximum temperature attained in Dublin is 23 degree which is greater than the maximum temperature attained in Cork(20.2 degree). Lowest temperature attained in Dublin is 0.4 degree which is much lower than the lowest temperature in Cork(4.7 degree). From this data it can be inferred that Dublin is overall hotter and relatively cooler than Cork for the month September in 2018. Average Rainfall in Cork(2.58) is higher than that of Dublin(1.46). Mean and Median of max temperature and minimum temperature in Dublin is almost same showing symmetric normal distribution.

##3.4: Create an S3 plot method for the class WeatherData that produces the following plots.##

```{r}
#Function definition (taking data and the main title of the plot as argument)
plot_ly.WeatherData <- function(Airport, TITLE){
  library(ggplot2)
  
  #Renaming Date column
  colnames(Airport)[colnames(Airport)=="ï..date"] <- "Date"
  
  #Initialising variables from input parameter for local use
  x  <- Airport$Date
  y1 <- Airport$maxtp
  y2 <- Airport$mintp
  y3 <- Airport$rain
  df <- data.frame(x,y1,y2)
  df2 <- data.frame(x, y3)
  
  require(ggplot2)
  library(plotly)
  
  #####################Layout Definitions#############################
  #Font
  f <- list(
    family = "Courier New, monospace",
    size = 18,
    color = "#7f7f7f"
  )
  
  #X-Axis
  x <- list(
    title = "Data for month September 18",
    titlefont = f
  )
  
  #Y-Axis of Top Graph
  ytop <- list(
    title = "Temperature in Degree",
    titlefont = f
  )
  
  #Y-Axis of bottom graph
  ybottom <- list(
    title = "Precipitation Amount",
    titlefont = f
  )
  
  #Values for indicators
  max <- Airport[which.max(Airport$maxtp), ]
  min <- Airport[which.min(Airport$mintp), ]
  highrain <- Airport[which.max(Airport$rain), ]
  rainx <- as.numeric(substr(x = highrain$Date, start = 1, stop = 2))-1
  
  #Annotation for max indicator
  a <- list(
    x = max$Date,
    y = max$maxtp,
    text = "Maximum Temp",
    xref = "x",
    yref = "y",
    showarrow = TRUE,
    # arrowhead = 7,
    ax = 20,
    ay = -40
  )
  
  #Annotation for min indicator
  b <- list(
    x = min$Date,
    y = min$mintp,
    text = "Minimum Temp",
    xref = "x",
    yref = "y",
    showarrow = TRUE,
    # arrowhead = 7,
    ax = 20,
    ay = -40
  )
  

  
  ###################Plotting graphs########################################
  
  #Plotting top graph
  p <- plot_ly(df, x = ~x) %>%
    # layout(title = TITLE) %>%
    add_trace(y = ~y1, name = 'max temperature', type='scatter', mode = 'lines', line = list(color = 'rgb(205, 12, 24)')) %>%
    add_trace(y = ~y2, name = 'min temperature', type='scatter',mode = 'lines', line = list(color = 'rgb(22, 96, 167)')) %>%
    layout(title = TITLE, xaxis = x, yaxis = ytop) %>%
    add_markers(x=x, y=y) %>%
    layout(annotations = a) %>%
    add_markers(x=x, y=y) %>%
    layout(annotations = b)
    #abline(v=c(1:30), col=c("grey")) (abline is not working for plotly but I am getting gray indicator lines on the graph. I have understood the functioning of abline. Have included it in comment to convey that I have learnt its usage but not used in the main code.)
  # print(p)
  
  
  #Plot bottom graph
  q <- plot_ly(y=y3, x=x, type='bar') %>%
    layout(xaxis = x, yaxis = ybottom) %>%
    add_segments(x = rainx, xend = rainx, y = 0, yend = highrain$rain, line = list(color = 'rgb(205, 12, 24)'))
  # print(q)
  subplot(p, q, nrows = 2, shareX = FALSE, shareY = FALSE, margin = 0.07)
  
}
```


```{r eval= FALSE}
#Testing the function for two airports
plot_ly.WeatherData(DublinAirport, "Dublin Airport Weather Data for Sept 2018")
plot_ly.WeatherData(CorkAirport, "Cork Airport Weather Data for Sept 2018")
```
